---
title: "Case1"
author: "Matt Chinchilla"
date: "8/27/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import Libraries
```{r}
library(tidyverse)
library(magrittr)

```

# Offline Data prep
Read in offline data and Parse data into a data frame called 'offline'
```{r}
# Read in the raw "offline" text file
txt = readLines('http://www.rdatasciencecases.org/Data/offline.final.trace.txt')

# Create a function to parse the data
processLine = function(x)
{
  tokens = strsplit(x, "[;=,]")[[1]]
  if (length(tokens) == 10) {
    return(NULL)
  }
  tmp = matrix(tokens[ - (1:10) ], , 4, byrow = TRUE)
  cbind(matrix(tokens[c(2, 4, 6:8, 10)], nrow(tmp), 6,
               byrow = TRUE), tmp)
}

lines = txt[ substr(txt, 1, 1) != "#" ]
tmp = lapply(lines, processLine)

# Convert the offline data to a data frame
offline = as.data.frame(do.call("rbind", tmp),stringsAsFactors = FALSE)

# Assign column names to the offline data frame
names(offline) = c("time", "scanMac", "posX", "posY", "posZ",
                   "orientation", "mac", "signal",
                   "channel", "type")

head(offline)

```



```{r}
# Observe one observation from the offline data frame
offline[offline$time==1139643118358,]
```

```{r}
# View the data frame exclusive of the scanMac, channel, and type columns
select(offline,-c(scanMac,channel,type))
```

```{r}
# See if any observations have a posZ-value not equal to 0.0
offline[offline['posZ']!='0.0',]
```

```{r}
# Get a full list of mac values and their types (we are interested in type == 3)
macTypeDF <- unique(select(offline, c(mac,type)))
macTypeDF[order(macTypeDF$type),]
```

```{r}
# List the unique mac values
vals = data.frame(table(offline['mac']))
vals[order(-vals$Freq),]

# Add the type to the an updated data frame
valsUpdated <- vals %>% inner_join(macTypeDF, by = c("Var1" = "mac"))
valsUpdated <- valsUpdated[order(valsUpdated$type, -valsUpdated$Freq),]

```

Refine offline DF to type 3 and mac addresses with the highest frequency
```{r}
#Change offline signal column from chr to integar
offline$signal %<>% as.integer

# Limit offline to type 3 observations
offline <- offline[offline$type != 1, ]

#These mac addresses are type 3 and have the most observations
keepMacs <- c('00:0f:a3:39:e1:c0',
              '00:0f:a3:39:dd:cd',
              '00:14:bf:b1:97:8a',
              '00:14:bf:3b:c7:c6',
              '00:14:bf:b1:97:90',
              '00:14:bf:b1:97:8d',
              '00:14:bf:b1:97:81'
              )

# Trim down offline DF to only the mac addresses that are type 3 with highest frequency.
offline <- offline[offline$mac %in% keepMacs ,]
```

Build final offline dataframe
```{r}
# Pivot the data frame (or cast it; make it wider) but putting the mac values and their associated signals in the columns
offlineOut<-select(offline, -c(channel,scanMac,type)) %>% pivot_wider(names_from = mac,values_from = signal, values_fn = list(signal=mean))

offlineOut$nas<-rowSums(is.na(offlineOut))

# View the final data frame
head(offlineOut)
```

# Online Data prep
read in Online data from website
```{r}
# Process the online data
onlineTxt = readLines("http://www.rdatasciencecases.org/Data/online.final.trace.txt")

onlineLines = onlineTxt[ substr(onlineTxt, 1, 1) != "#" ]
onlineTmp = lapply(onlineLines, processLine)

# Convert the online data to a data frame
online = as.data.frame(do.call("rbind", onlineTmp),stringsAsFactors = FALSE)

head(online)

names(online) = c("time", "scanMac", "posX", "posY", "posZ",
                   "orientation", "mac", "signal",
                   "channel", "type")

head(online)
```

```{r}
# View the data frame exclusive of the scanMac, channel, and type columns
select(online,-c(scanMac,channel,type))
```

```{r}
# See if any observations have a posZ-value not equal to 0.0
online[online['posZ']!='0.0',]

```

```{r}
# List the unique mac values and their frequencies in the offline data frame
onlineVals = data.frame(table(online['mac']))
onlineVals[order(-onlineVals$Freq),]

online$signal %<>% as.integer

online <- online[online$mac %in% keepMacs ,]
```
Create final online dataframe
```{r}
# Pivot the data frame (or cast it; make it wider) but putting the mac values and their associated signals in the columns
onlineOut<-select(online, -c(channel,scanMac,type)) %>% pivot_wider(names_from = mac,values_from = signal, values_fn = list(signal=mean))

onlineOut$nas<-rowSums(is.na(onlineOut))

# View the final data frame
head(onlineOut)
```

Add posXY column to both offlineOut and OnlineOut dataframes
```{r}
offlineOut$posXY <- paste(offlineOut$posX, offlineOut$posY, sep="-")
length(unique(offlineOut$posXY))

onlineOut$posXY <- paste(onlineOut$posX, onlineOut$posY, sep="-")
length(unique(onlineOut$posXY))

```

Remove N/A values and remove posZ column from data frames
```{R}
<<<<<<< HEAD
=======
#Remove NA's
offlineOut <- na.omit(offlineOut)
onlineOut <- na.omit(onlineOut)
>>>>>>> ec5052f80c763782ff73bb4e72670a4d29d7af38

#Remove posZ column
offlineOut <- offlineOut %>% subset(select = -c(posZ))
onlineOut <- onlineOut %>% subset(select = -c(posZ))

```


Convert posX,posY, and orientation columns from chr to num and add angle column to OfflineOut and OnlineOut Dataframes
```{r}
# Convert char columns to numeric
#Offline
offlineOut$orientation <- as.numeric(offlineOut$orientation)
offlineOut$posX <- as.numeric(offlineOut$posX)
offlineOut$posY <- as.numeric(offlineOut$posY)

#Online
onlineOut$orientation <- as.numeric(onlineOut$orientation)
onlineOut$posX <- as.numeric(onlineOut$posX)
onlineOut$posY <- as.numeric(onlineOut$posY)

# create a function that will round off to the nearest major angle
roundOrientation = function(angles) {
  refs = seq(0, by = 45, length  = 9)
  q = sapply(angles, function(o) which.min(abs(o - refs)))
  c(refs[1:8], 0)[q]
}

# we built it, now apply it
offlineOut$angle = roundOrientation(offlineOut$orientation)
onlineOut$angle = roundOrientation(onlineOut$orientation)
```

<<<<<<< HEAD
```{r}
offline0df = offlineOut[offlineOut$angle == 0,]
offline45df = offlineOut[offlineOut$angle == 45,]
offline90df = offlineOut[offlineOut$angle == 90,]
offline135df = offlineOut[offlineOut$angle == 135,]
offline180df = offlineOut[offlineOut$angle == 180,]
offline225df = offlineOut[offlineOut$angle == 225,]
offline270df = offlineOut[offlineOut$angle == 270,]
offline315df = offlineOut[offlineOut$angle == 315,]

online0df = onlineOut[onlineOut$angle == 0,]
online45df = onlineOut[onlineOut$angle == 45,]
online90df = onlineOut[onlineOut$angle == 90,]
online135df = onlineOut[onlineOut$angle == 135,]
online180df = onlineOut[onlineOut$angle == 180,]
online225df = onlineOut[onlineOut$angle == 225,]
online270df = onlineOut[onlineOut$angle == 270,]
online315df = onlineOut[onlineOut$angle == 315,]

offlinedflist = list(offline0df,offline45df,offline90df,
                    offline135df,offline180df,offline225df,
                    offline270df,offline315df)

onlinedflist = list(online0df,online45df,online90df,
                    online135df,online180df,online225df,
                    online270df,online315df)

#Remove NA's
for (i in 1:8) {
  for (j in colnames(offline0df[5:11])) {
    offlinedflist[[i]][[j]][is.na(offlinedflist[[i]][[j]])] <- mean(na.omit(offlinedflist[[i]][[j]]))
    onlinedflist[[i]][[j]][is.na(onlinedflist[[i]][[j]])] <- mean(na.omit(offlinedflist[[i]][[j]]))
  }
}
offline0df = data.frame(offlinedflist[1])
offline45df = data.frame(offlinedflist[2])
offline90df = data.frame(offlinedflist[3])
offline135df = data.frame(offlinedflist[4])
offline180df = data.frame(offlinedflist[5])
offline225df = data.frame(offlinedflist[6])
offline270df = data.frame(offlinedflist[7])
offline315df = data.frame(offlinedflist[8])

online0df = data.frame(onlinedflist[1])
online45df = data.frame(onlinedflist[2])
online90df = data.frame(onlinedflist[3])
online135df = data.frame(onlinedflist[4])
online180df = data.frame(onlinedflist[5])
online225df = data.frame(onlinedflist[6])
online270df = data.frame(onlinedflist[7])
online315df = data.frame(onlinedflist[8])

sum(is.na(offline0df))
sum(is.na(offlinedflist[2]))



```

```{r}
gpOffline0df <- offline0df %>% group_by(posX, posY, angle) %>%
  summarize(
            '00.14.bf.b1.97.8a' = mean(X00.14.bf.b1.97.8a),
            '00.14.bf.b1.97.90' = mean(X00.14.bf.b1.97.90),
            '00.0f.a3.39.e1.c0' = mean(X00.0f.a3.39.e1.c0),
            '00.14.bf.b1.97.8d' = mean(X00.14.bf.b1.97.8d),
            '00.14.bf.b1.97.81' = mean(X00.14.bf.b1.97.81),
            '00.14.bf.3b.c7.c6' = mean(X00.14.bf.3b.c7.c6),
            '00.0f.a3.39.dd.cd' = mean(X00.0f.a3.39.dd.cd))
gpOffline45df <- offline45df %>% group_by(posX, posY, angle) %>%
  summarize(
            '00.14.bf.b1.97.8a' = mean(X00.14.bf.b1.97.8a),
            '00.14.bf.b1.97.90' = mean(X00.14.bf.b1.97.90),
            '00.0f.a3.39.e1.c0' = mean(X00.0f.a3.39.e1.c0),
            '00.14.bf.b1.97.8d' = mean(X00.14.bf.b1.97.8d),
            '00.14.bf.b1.97.81' = mean(X00.14.bf.b1.97.81),
            '00.14.bf.3b.c7.c6' = mean(X00.14.bf.3b.c7.c6),
            '00.0f.a3.39.dd.cd' = mean(X00.0f.a3.39.dd.cd))
gpOffline90df <- offline90df %>% group_by(posX, posY, angle) %>%
  summarize(
            '00.14.bf.b1.97.8a' = mean(X00.14.bf.b1.97.8a),
            '00.14.bf.b1.97.90' = mean(X00.14.bf.b1.97.90),
            '00.0f.a3.39.e1.c0' = mean(X00.0f.a3.39.e1.c0),
            '00.14.bf.b1.97.8d' = mean(X00.14.bf.b1.97.8d),
            '00.14.bf.b1.97.81' = mean(X00.14.bf.b1.97.81),
            '00.14.bf.3b.c7.c6' = mean(X00.14.bf.3b.c7.c6),
            '00.0f.a3.39.dd.cd' = mean(X00.0f.a3.39.dd.cd))
gpOffline90df <- offline90df %>% group_by(posX, posY, angle) %>%
  summarize(
            '00.14.bf.b1.97.8a' = mean(X00.14.bf.b1.97.8a),
            '00.14.bf.b1.97.90' = mean(X00.14.bf.b1.97.90),
            '00.0f.a3.39.e1.c0' = mean(X00.0f.a3.39.e1.c0),
            '00.14.bf.b1.97.8d' = mean(X00.14.bf.b1.97.8d),
            '00.14.bf.b1.97.81' = mean(X00.14.bf.b1.97.81),
            '00.14.bf.3b.c7.c6' = mean(X00.14.bf.3b.c7.c6),
            '00.0f.a3.39.dd.cd' = mean(X00.0f.a3.39.dd.cd))
gpOffline135df <- offline135df %>% group_by(posX, posY, angle) %>%
  summarize(
            '00.14.bf.b1.97.8a' = mean(X00.14.bf.b1.97.8a),
            '00.14.bf.b1.97.90' = mean(X00.14.bf.b1.97.90),
            '00.0f.a3.39.e1.c0' = mean(X00.0f.a3.39.e1.c0),
            '00.14.bf.b1.97.8d' = mean(X00.14.bf.b1.97.8d),
            '00.14.bf.b1.97.81' = mean(X00.14.bf.b1.97.81),
            '00.14.bf.3b.c7.c6' = mean(X00.14.bf.3b.c7.c6),
            '00.0f.a3.39.dd.cd' = mean(X00.0f.a3.39.dd.cd))
gpOffline180df <- offline180df %>% group_by(posX, posY, angle) %>%
  summarize(
            '00.14.bf.b1.97.8a' = mean(X00.14.bf.b1.97.8a),
            '00.14.bf.b1.97.90' = mean(X00.14.bf.b1.97.90),
            '00.0f.a3.39.e1.c0' = mean(X00.0f.a3.39.e1.c0),
            '00.14.bf.b1.97.8d' = mean(X00.14.bf.b1.97.8d),
            '00.14.bf.b1.97.81' = mean(X00.14.bf.b1.97.81),
            '00.14.bf.3b.c7.c6' = mean(X00.14.bf.3b.c7.c6),
            '00.0f.a3.39.dd.cd' = mean(X00.0f.a3.39.dd.cd))
gpOffline225df <- offline45df %>% group_by(posX, posY, angle) %>%
  summarize(
            '00.14.bf.b1.97.8a' = mean(X00.14.bf.b1.97.8a),
            '00.14.bf.b1.97.90' = mean(X00.14.bf.b1.97.90),
            '00.0f.a3.39.e1.c0' = mean(X00.0f.a3.39.e1.c0),
            '00.14.bf.b1.97.8d' = mean(X00.14.bf.b1.97.8d),
            '00.14.bf.b1.97.81' = mean(X00.14.bf.b1.97.81),
            '00.14.bf.3b.c7.c6' = mean(X00.14.bf.3b.c7.c6),
            '00.0f.a3.39.dd.cd' = mean(X00.0f.a3.39.dd.cd))
gpOffline270df <- offline45df %>% group_by(posX, posY, angle) %>%
  summarize(
            '00.14.bf.b1.97.8a' = mean(X00.14.bf.b1.97.8a),
            '00.14.bf.b1.97.90' = mean(X00.14.bf.b1.97.90),
            '00.0f.a3.39.e1.c0' = mean(X00.0f.a3.39.e1.c0),
            '00.14.bf.b1.97.8d' = mean(X00.14.bf.b1.97.8d),
            '00.14.bf.b1.97.81' = mean(X00.14.bf.b1.97.81),
            '00.14.bf.3b.c7.c6' = mean(X00.14.bf.3b.c7.c6),
            '00.0f.a3.39.dd.cd' = mean(X00.0f.a3.39.dd.cd))
gpOffline315df <- offline45df %>% group_by(posX, posY, angle) %>%
  summarize(
            '00.14.bf.b1.97.8a' = mean(X00.14.bf.b1.97.8a),
            '00.14.bf.b1.97.90' = mean(X00.14.bf.b1.97.90),
            '00.0f.a3.39.e1.c0' = mean(X00.0f.a3.39.e1.c0),
            '00.14.bf.b1.97.8d' = mean(X00.14.bf.b1.97.8d),
            '00.14.bf.b1.97.81' = mean(X00.14.bf.b1.97.81),
            '00.14.bf.3b.c7.c6' = mean(X00.14.bf.3b.c7.c6),
            '00.0f.a3.39.dd.cd' = mean(X00.0f.a3.39.dd.cd))
```

```{r}
gpOffline0df
online0df



```
=======

>>>>>>> ec5052f80c763782ff73bb4e72670a4d29d7af38
