---
title: "Spam detector"
author: "Matt Chinchilla, Rikel Djoko, Drew Larsen"
date: "9/29/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(XML)
library(stringr)
library(dplyr)
library(tidyverse)
library(corrplot)
library(rpart)
```


### Exploratory Data Analysis:

```{r, include=True}
setwd("C:/Users/radjoko/Downloads/DS 7333-QTW/week 5-6")
getwd()
#load the rda file
data <- load(file = "data.Rda")
head(data)
head(emailDFrp)
```
### Data Description

The dataset is made up of a corpus of emails from SpamAssassin.org (Apache 4). In total, there are 9348 unique emails. It contains 29 predictor variables and one response variable named isSpam. Of the 30 total variables, 17 are boolean factor variables and the remaining 13 variables are numeric variables. Our goal is to the existing email previously classified  to classify future incoming email based on the value of the predictor variable. See table below for the list of the variable
```{r, include=True}
### data dimension description
str(emailDFrp)
```


```{r, include=True}
### data dimension
dim(emailDFrp)
```




```{r, include=True}
### summary statistic ####
summary(emailDFrp)
```
```{r, include=True}
### data feature
colnames(emailDFrp)
```
```{r, include=True}
### data dimension description
str(emailDFrp)
```

```{r, include=True}
### Checking for NA value
print(paste0("number of missing value: ", sum(is.na(emailDFrp))))
```
The evaluation of the dataset found 357  missing observations.Rather than doing a preliminary imputation we will let the random forest method to take care of those missing numeric and categorical predictor variables.
```{r, include=True}
###Plot distribution of email vs spam
ggplot(emailDFrp, aes(x=isSpam)) +
  geom_bar(fill = "#0073C2FF")+
  ggtitle("Figure 1: Spam vs. Non-Spam Split")
```
```{r, include=True}
#### plot relationship with spam  and other categorical variable
ggplot(data = emailDFrp) +
  geom_count(mapping = aes(x = isSpam, y = isRe))+
  ggtitle("Spam vs. isRe count")
```
```{r, include=True}
#### plot relationship with predictor and other categorical variable
ggplot(data = emailDFrp) +
  geom_count(mapping = aes(x = isSpam, y = priority))+
  ggtitle("Spam vs. mail priority count")
```
```{r, include=True}
# Count Observations by Two Groups
emailDFrp  %>%
  count("priority", "isSpam")
```

```{r, include=True}
# #pair count correlation
emailDFrp %>% 
  count(isInReplyTo, isSpam) %>%  
  ggplot(mapping = aes(x = isInReplyTo, y = isSpam)) +
    geom_tile(mapping = aes(fill = n))
```

```{r, include=True}
 # #pair count correlation
emailDFrp %>% 
  count(isRe, isSpam) %>%  
  ggplot(mapping = aes(x = isRe, y = isSpam)) +
    geom_tile(mapping = aes(fill = n))

```

######################################## building tree####
```{r, include=True}
# split variable data
 set.seed(2568)
 n <- nrow(emailDFrp)
 train <- sort(sample(1:n, floor(n/2)))
 train_df <- emailDFrp[train,]
 
 # negate the indices to get the test data:
 test_df <- emailDFrp[-train,]

```

```{r, include=True}
# build rpart tree see parameter description:  #https://www.rdocumentation.org/packages/rpart/versions/4.1-15/topics/rpart.control
# example parameter: 

email.rp <-rpart(isSpam ~ . ,                         # formula: . means all
                 data = emailDFrp,                       # data frame
                 subset = train,                       # indices of training set
                 method = "class",                     # classification tree
                 parms = list(split = "information"),  # use entropy/deviance
                 usesurrogate = 2,                     # since having  missing values
                 maxsurrogate = 1,  
                 cp = 0.1,                               # no size penalty
                 minsplit = 5,                         # Nodes of size 5 (or 
                 # more) can be split,
                 minbucket = 2,                        # provided each sub-node
                 # contains at least 2 obs.
 )
 summary(email.rp)
```
```{r, include=True}
# plot tree
 #newwindow()
 plot(email.rp, 
      uniform=TRUE,
      compress=TRUE,
      margin = .2)
 text(email.rp, 
      use.n=TRUE, 
      all = TRUE,
      fancy = TRUE)

```

```{r, include=True}
 # plot tree https://www.gormanalysis.com/blog/decision-trees-in-r-using-rpart/
 #newwindow()
 library(rattle)
 library(rpart.plot)
 library(RColorBrewer)
 
 # plot tree
 fancyRpartPlot(email.rp, caption = NULL)

```

```{r, include=True}
 # Tree acccuracy (confusion matrix)
predict_validation<- predict(email.rp, newdata = test_df)
m_ct<-confusionMatrix(predict_validation,test_df$isSpam)
m_ct$cm_ct$overall['Accuracy']
```
